Protected Areas Practical

### 2024-11-22 tutorial/MSc_PA_practical_excercise.Rmd
loading rasters

# Why systematic conservation planning?

To understand the value of using an optimization, let's first try to solve [this problem](https://marxansolutions.org/training/#:~:text=Conservation%20Planning%20Exercise,with%20the%20lowest%20possible%20cost.) by selecting cells for conservation to meet targets for 3 species for the least cost. It's not as easy as it seems.

# Installation
## R, Rstudio and Rtools

Make sure you have Rstudio, R and Rtools installed:

- R: https://cran.r-project.org/bin/windows/base/
- Rtools: https://cran.r-project.org/bin/windows/Rtools/
- Rstudio: https://posit.co/download/rstudio-desktop/

@@

Run the following code to ensure all packages are installed:

- Packages for making pretty Rmarkdown documents:
```r
install.packages("knitr")
install.packages("rmdformats")
install.packages("prettydoc")
```
- Packages for protected areas:

```r 
install.packages("prioritizr")
install.packages("prioritizrdata")
install.packages("terra")
install.packages("highs", repos = "https://cran.rstudio.com/")
```
# Load Packages

@@

## Specify the directory that you are working from

Make sure you change the code to indicate your directory, and make sure that you use forward slash `/` or double back slash `\\` to separate your directories.

```{r dir}
basepath = "C:/Users/kdh10kg/OneDrive - The Royal Botanic Gardens, Kew/teaching/2024-2025/MSC_practical/"
```

@@

Each layer describes the spatial distribution of a feature. Here, our feature data correspond to different plant species. Specifically, the cell values denote the relative probability of a species occuring, with higher values indicating higher probability.

>>>>

# Installation
## R and Rstudio

Make sure you have Rstudio and R installed:

- R: https://cran.r-project.org/bin/windows/base/
- Rstudio: https://posit.co/download/rstudio-desktop/

## Download the practical data

1. Go to https://github.com/KiranLDA/MSc_Protected_Areas
2. Click on the big green button 'CODE'.
3. Select 'Download ZIP'.
4. Save the file to your desktop.
5. Once opened unzip/extract all files in the ZIP file.

@@

Go to `code/01_installation.R` to get a list of R packages to install. They are also copied below:

```r 
install.packages("prioritizr")
install.packages("prioritizrdata")
install.packages("terra")
install.packages("highs", repos = "https://cran.rstudio.com/")
```

# Excercise 1

While the packages install, we can do this exercise to understand the value of using an optimization. Let's first try to solve [this problem](https://marxansolutions.org/training/#:~:text=Conservation%20Planning%20Exercise,with%20the%20lowest%20possible%20cost.) by selecting cells for conservation to meet targets for 3 species for the least cost. Once you meet the target, you can compare your solution with the optimal solution.

# Load Packages

@@

@@

Each layer describes the spatial distribution of a feature. Here, our feature data correspond to different plant species. Specifically, if the cell (or planning unit) has a 1, as species is predicted to be present in it based on the SDM. If the cell has a 0, then it is predicted to be absent.
